---
title: "p8105_hw3_at3535"
author: "Amanda Tsai"
date: "10/10/2020"
output: github_document
---


```{r setup}
library(tidyverse)
library(p8105.datasets)
data("instacart")
knitr::opts_chunk$set(
  fig.width = 10,
  fig.asp = .6,
  out.width = "90%"
)
```

# Problem 1


# Problem 2 

The following code chunk loads the accelerometer dataset and cleans the dataset by encoding with reasonable variable classes, adds a new variable week_weekend that determines whether the day observed is a weekday or weekend, and combines all activity observations into one variable column. 

```{r accel}
accel_df = 
  read_csv(file = "./data/accel_data.csv") %>%
  janitor::clean_names() %>%
  mutate(day = as.factor(day)) %>%
  mutate(week_weekend = as.factor(ifelse(day %in% c("Saturday", "Sunday"), "weekend", "weekday"))) %>%
  pivot_longer(activity_1:activity_1440,
               names_to = "minute",
               names_prefix = "activity_",
               values_to = "activity_count") %>%
  mutate(
    minute = as.integer(minute),
    week = as.factor(week), 
    day_id = as.factor(day_id)) 

accel_df
skimr::skim(accel_df)
```

The resulting dataset has the existing variables: **`r colnames(accel_df)`** which respectively are of types:  **`r map(accel_df, class)`**.

The dimensions of the dataset are `r nrow(accel_df)` x `r ncol(accel_df)`, meaning that there are a total of `r nrow(accel_df)` observations for `r ncol(accel_df)` different variables. There are 35 unique entries for the day_id variable, showing that the dataset contains observations over a span of 35 days.

The key variables in this dataset are **day, minute** and **activity_count**. 

```{r aggregate}
total_activity = 
  accel_df %>% 
  group_by(day_id) %>%
  summarize(daily_activity = sum(activity_count))

total_activity
skimr::skim(total_activity)
```

The trends are not quite clear. 

```{r accel_plot}
accel_df %>%
  ggplot(aes(x = factor(minute), y = activity_count)) +
  geom_point(alpha = 0.5, aes(color = day)) +
  scale_x_discrete(breaks = seq(from = 0, to = 1440, by = 60), 
                   labels = seq(from = 0, to = 1440, by = 60) ) +
  geom_smooth(se = FALSE)


```


# Problem 3 
```{r setup_noaa}
data("ny_noaa")
```





